---
title: "A1 APM466"
author: "Ingrid Leung"
date: "2023-02-02"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
setwd("/Users/ingridleung/Downloads")
bond_data <- read.csv("APM466 A1 Coupons - Sheet1.csv")
library("jrvFinance")
library("anytime")
```


Coupon Function
```{r}
coupons <- vector()
coupon_rates <- c(bond_data$Coupon)
for (t in 1:length(coupon_rates)) {
  t1 <- coupon_rates[t]*100*1/2
  coupons[t] <- t1
}
coupons

coupon_payments <- as.numeric(unlist(coupons))
```

YTM
```{r}
Dates <- c('01-16-2023', '01-17-2023', '01-18-2023', '01-19-2023', '01-20-2023', '01-23-2023', '01-24-2023', '01-25-2023', '01-26-2023', '01-27-2023')

Dates <- anydate(Dates)
Prices <- bond_data[c("Jan.16", "Jan.17", "Jan.18", "Jan.19", "Jan.20", "Jan.23", "Jan.24", "Jan.25", "Jan.26", "Jan.27")]

YTM <- matrix(,ncol=10,nrow=10)

```

YTM Jan 16
```{r}
maturity_dates <- anydate(bond_data$Maturity.Date)
ytm_Jan16 <- c()
for (x in 1:10){
  yield1 <- bond.yield(settle="2023-01-16", mature=maturity_dates[x], coupon=bond_data$Coupon[x], freq=2, price=bond_data$Jan.16[x])
  ytm_Jan16 <- append(ytm_Jan16, yield1)
}
ytm_Jan16
```

YTM Jan 17
```{r}
ytm_Jan17 <- c()
for (x in 1:10){
  yield2 <- bond.yield(settle="2023-01-17", mature=maturity_dates[x], coupon=bond_data$Coupon[x], freq=2, price=bond_data$Jan.17[x])
  ytm_Jan17 <- append(ytm_Jan17, yield2)
}
ytm_Jan17
```

YTM Jan 18
```{r}
ytm_Jan18 <- c()
for (x in 1:10){
  yield3 <- bond.yield(settle="2023-01-18", mature=maturity_dates[x], coupon=bond_data$Coupon[x], freq=2, price=bond_data$Jan.18[x])
  ytm_Jan18 <- append(ytm_Jan18, yield3)
}
ytm_Jan18
```

YTM Jan 19
```{r}
ytm_Jan19 <- c()
for (x in 1:10){
  yield4 <- bond.yield(settle="2023-01-19", mature=maturity_dates[x], coupon=bond_data$Coupon[x], freq=2, price=bond_data$Jan.19[x])
  ytm_Jan19 <- append(ytm_Jan19, yield4)
}
ytm_Jan19
```

YTM Jan 20
```{r}
ytm_Jan20 <- c()
for (x in 1:10){
  yield5 <- bond.yield(settle="2023-01-20", mature=maturity_dates[x], coupon=bond_data$Coupon[x], freq=2, price=bond_data$Jan.20[x])
  ytm_Jan20 <- append(ytm_Jan20, yield5)
}
ytm_Jan20
```

YTM Jan 23
```{r}
ytm_Jan23 <- c()
for (x in 1:10){
  yield6 <- bond.yield(settle="2023-01-23", mature=maturity_dates[x], coupon=bond_data$Coupon[x], freq=2, price=bond_data$Jan.23[x])
  ytm_Jan23 <- append(ytm_Jan23, yield6)
}
ytm_Jan23
```

YTM Jan 24
```{r}
ytm_Jan24 <- c()
for (x in 1:10){
  yield7 <- bond.yield(settle="2023-01-24", mature=maturity_dates[x], coupon=bond_data$Coupon[x], freq=2, price=bond_data$Jan.24[x])
  ytm_Jan24 <- append(ytm_Jan24, yield7)
}
ytm_Jan24
```

YTM Jan 25
```{r}
ytm_Jan25 <- c()
for (x in 1:10){
  yield8 <- bond.yield(settle="2023-01-25", mature=maturity_dates[x], coupon=bond_data$Coupon[x], freq=2, price=bond_data$Jan.25[x])
  ytm_Jan25 <- append(ytm_Jan25, yield8)
}
ytm_Jan25
```

YTM Jan 26
```{r}
ytm_Jan26 <- c()
for (x in 1:10){
  yield9 <- bond.yield(settle="2023-01-26", mature=maturity_dates[x], coupon=bond_data$Coupon[x], freq=2, price=bond_data$Jan.26[x])
  ytm_Jan26 <- append(ytm_Jan26, yield9)
}
ytm_Jan26
```

YTM Jan 27
```{r}
ytm_Jan27 <- c()
for (x in 1:10){
  yield10 <- bond.yield(settle="2023-01-27", mature=maturity_dates[x], coupon=bond_data$Coupon[x], freq=2, price=bond_data$Jan.27[x])
  ytm_Jan27 <- append(ytm_Jan27, yield10)
}
ytm_Jan27
```

Plot YTM
```{r}
q <- c(0.5, 1, 1.5, 2, 2.5, 3, 3.5, 4, 4.5, 5)
plot(q, ytm_Jan16, type='l', main="5 Year Yield Curve", xlab='Year', ylab = 'ytm (%)', ylim=c(0.025, 0.05), col =1)
lines(q, ytm_Jan17, col=2)
lines(q, ytm_Jan18, col=3)
lines(q, ytm_Jan19, col=4)
lines(q, ytm_Jan20, col=5)
lines(q, ytm_Jan23, col=6)
lines(q, ytm_Jan24, col=7)
lines(q, ytm_Jan25, col=8)
lines(q, ytm_Jan26, col=9)
lines(q, ytm_Jan27, col=10)
legend("topright", legend=Dates, cex=0.75, col=1:10, lty=1)
```

Spot Rates for Jan 16
```{r}
spot_rates_Jan16 <- c()
CAN1.75Mar1_Jan16_SR <- (-log(99.71/100.87))*8
spot_rates_Jan16 <- append(spot_rates_Jan16, CAN1.75Mar1_Jan16_SR)

time_until_maturity <- c(9/12, 13/12, 19/12, 25/12, 31/12, 37/12, 43/12, 49/12, 55/12)

discounted = 0
LHS = 0
for (j in 2:length(coupon_payments)){
    LHS = coupon_payments[j] * exp(-spot_rates_Jan16[j-1] * time_until_maturity[j-1])
    discounted = bond_data$Jan.16[j] - LHS
    spot_rates_Jan16 <- append(spot_rates_Jan16, (-log(discounted/(100+coupon_payments[j])))/time_until_maturity[j])
}

#Last one done manually
j <- coupon_payments[10] * exp(-spot_rates_Jan16[9] * time_until_maturity[9])
k <- bond_data$Jan.16[10] - j
spot_rates_Jan16[10] <- -log(k/(100+coupon_payments[10]))/time_until_maturity[9]

spot_rates_Jan16
```

Spot Rates for Jan 17
```{r}
spot_rates_Jan17 <- c()
CAN1.75Mar1_Jan17_SR <- (-log(99.7/100.87))*8
spot_rates_Jan17 <- append(spot_rates_Jan17, CAN1.75Mar1_Jan17_SR)

discounted1 = 0
LHS1 = 0
for (j in 2:length(coupon_payments)){
    LHS1 = coupon_payments[j] * exp(-spot_rates_Jan17[j-1] * time_until_maturity[j])
    discounted1 = bond_data$Jan.17[j] - LHS1
    spot_rates_Jan17 <- append(spot_rates_Jan17, (-log(discounted1/(100+coupon_payments[j])))/time_until_maturity[j])
}

#Last one done manually
j1 <- coupon_payments[10] * exp(-spot_rates_Jan17[9] * time_until_maturity[9])
k1 <- bond_data$Jan.17[10] - j1
spot_rates_Jan17[10] <- -log(k1/(100+coupon_payments[10]))/time_until_maturity[9]

spot_rates_Jan17
```

Spot Rates for Jan 18
```{r}
spot_rates_Jan18 <- c()
CAN1.75Mar1_Jan18_SR <- (-log(99.72/100.87))*8
spot_rates_Jan18 <- append(spot_rates_Jan18, CAN1.75Mar1_Jan18_SR)

discounted2 = 0
LHS2 = 0
for (j in 2:length(coupon_payments)){
    LHS2 = coupon_payments[j] * exp(-spot_rates_Jan18[j-1] * time_until_maturity[j])
    discounted2 = bond_data$Jan.18[j] - LHS2
    spot_rates_Jan18 <- append(spot_rates_Jan18, (-log(discounted2/(100+coupon_payments[j])))/time_until_maturity[j])
}

#Last one done manually
j2 <- coupon_payments[10] * exp(-spot_rates_Jan18[9] * time_until_maturity[9])
k2 <- bond_data$Jan.18[10] - j2
spot_rates_Jan18[10] <- -log(k2/(100+coupon_payments[10]))/time_until_maturity[9]

spot_rates_Jan18
```

Spot Rates for Jan 19
```{r}
spot_rates_Jan19 <- c()
CAN1.75Mar1_Jan19_SR <- (-log(99.72/100.87))*8
spot_rates_Jan19 <- append(spot_rates_Jan19, CAN1.75Mar1_Jan19_SR)

discounted3 = 0
LHS3 = 0
for (j in 2:length(coupon_payments)){
    LHS3 = coupon_payments[j] * exp(-spot_rates_Jan19[j-1] * time_until_maturity[j])
    discounted3 = bond_data$Jan.19[j] - LHS3
    spot_rates_Jan19 <- append(spot_rates_Jan19, (-log(discounted3/(100+coupon_payments[j])))/time_until_maturity[j])
}

#Last one done manually
j3 <- coupon_payments[10] * exp(-spot_rates_Jan19[9] * time_until_maturity[9])
k3 <- bond_data$Jan.19[10] - j3
spot_rates_Jan19[10] <- -log(k3/(100+coupon_payments[10]))/time_until_maturity[9]

spot_rates_Jan19
```

Spot Rates for Jan 20
```{r}
spot_rates_Jan20 <- c()
CAN1.75Mar1_Jan20_SR <- (-log(99.74/100.87))*8
spot_rates_Jan20 <- append(spot_rates_Jan20, CAN1.75Mar1_Jan20_SR)

discounted4 = 0
LHS4 = 0
for (j in 2:length(coupon_payments)){
    LHS4 = coupon_payments[j] * exp(-spot_rates_Jan20[j-1] * time_until_maturity[j])
    discounted4 = bond_data$Jan.20[j] - LHS4
    spot_rates_Jan20 <- append(spot_rates_Jan20, (-log(discounted4/(100+coupon_payments[j])))/time_until_maturity[j])
}

#Last one done manually
j4 <- coupon_payments[10] * exp(-spot_rates_Jan20[9] * time_until_maturity[9])
k4 <- bond_data$Jan.20[10] - j4
spot_rates_Jan20[10] <- -log(k4/(100+coupon_payments[10]))/time_until_maturity[9]

spot_rates_Jan20
```

Spot Rates for Jan 23
```{r}
spot_rates_Jan23 <- c()
CAN1.75Mar1_Jan23_SR <- (-log(99.75/100.87))*8
spot_rates_Jan23 <- append(spot_rates_Jan23, CAN1.75Mar1_Jan23_SR)

discounted5 = 0
LHS5 = 0
for (j in 2:length(coupon_payments)){
    LHS5 = coupon_payments[j] * exp(-spot_rates_Jan23[j-1] * time_until_maturity[j])
    discounted5 = bond_data$Jan.23[j] - LHS5
    spot_rates_Jan23 <- append(spot_rates_Jan23, (-log(discounted5/(100+coupon_payments[j])))/time_until_maturity[j])
}

#Last one done manually
j5 <- coupon_payments[10] * exp(-spot_rates_Jan23[9] * time_until_maturity[9])
k5 <- bond_data$Jan.23[10] - j5
spot_rates_Jan23[10] <- -log(k5/(100+coupon_payments[10]))/time_until_maturity[9]

spot_rates_Jan23
```

Spot Rates for Jan 24
```{r}
spot_rates_Jan24 <- c()
CAN1.75Mar1_Jan24_SR <- (-log(99.76/100.87))*8
spot_rates_Jan24 <- append(spot_rates_Jan24, CAN1.75Mar1_Jan24_SR)

discounted6 = 0
LHS6 = 0
for (j in 2:length(coupon_payments)){
    LHS6 = coupon_payments[j] * exp(-spot_rates_Jan24[j-1] * time_until_maturity[j])
    discounted6 = bond_data$Jan.24[j] - LHS6
    spot_rates_Jan24 <- append(spot_rates_Jan24, (-log(discounted6/(100+coupon_payments[j])))/time_until_maturity[j])
}

#Last one done manually
j6 <- coupon_payments[10] * exp(-spot_rates_Jan24[9] * time_until_maturity[9])
k6 <- bond_data$Jan.24[10] - j6
spot_rates_Jan24[10] <- -log(k6/(100+coupon_payments[10]))/time_until_maturity[9]

spot_rates_Jan24
```

Spot Rates for Jan 25
```{r}
spot_rates_Jan25 <- c()
CAN1.75Mar1_Jan25_SR <- (-log(99.77/100.87))*8
spot_rates_Jan25 <- append(spot_rates_Jan25, CAN1.75Mar1_Jan25_SR)

discounted7 = 0
LHS7 = 0
for (j in 2:length(coupon_payments)){
    LHS7 = coupon_payments[j] * exp(-spot_rates_Jan25[j-1] * time_until_maturity[j])
    discounted7 = bond_data$Jan.25[j] - LHS7
    spot_rates_Jan25 <- append(spot_rates_Jan25, (-log(discounted7/(100+coupon_payments[j])))/time_until_maturity[j])
}

#Last one done manually
j7 <- coupon_payments[10] * exp(-spot_rates_Jan25[9] * time_until_maturity[9])
k7 <- bond_data$Jan.25[10] - j7
spot_rates_Jan25[10] <- -log(k7/(100+coupon_payments[10]))/time_until_maturity[9]

spot_rates_Jan25
```

Spot Rates for Jan 26
```{r}
spot_rates_Jan26 <- c()
CAN1.75Mar1_Jan26_SR <- (-log(99.78/100.87))*8
spot_rates_Jan26 <- append(spot_rates_Jan26, CAN1.75Mar1_Jan26_SR)

discounted8 = 0
LHS8 = 0
for (j in 2:length(coupon_payments)){
    LHS8 = coupon_payments[j] * exp(-spot_rates_Jan26[j-1] * time_until_maturity[j])
    discounted8 = bond_data$Jan.26[j] - LHS8
    spot_rates_Jan26 <- append(spot_rates_Jan26, (-log(discounted8/(100+coupon_payments[j])))/time_until_maturity[j])
}
#Last one done manually
j8 <- coupon_payments[10] * exp(-spot_rates_Jan26[9] * time_until_maturity[9])
k8 <- bond_data$Jan.26[10] - j8
spot_rates_Jan26[10] <- -log(k8/(100+coupon_payments[10]))/time_until_maturity[9]

spot_rates_Jan26
```

Spot Rates for Jan 27
```{r}
spot_rates_Jan27 <- c()
CAN1.75Mar1_Jan27_SR <- (-log(99.79/100.87))*8
spot_rates_Jan27 <- append(spot_rates_Jan27, CAN1.75Mar1_Jan27_SR)

discounted9 = 0
LHS9 = 0
for (j in 2:length(coupon_payments)){
    LHS9 = coupon_payments[j] * exp(-spot_rates_Jan27[j-1] * time_until_maturity[j])
    discounted9 = bond_data$Jan.27[j] - LHS9
    spot_rates_Jan27 <- append(spot_rates_Jan27, (-log(discounted9/(100+coupon_payments[j])))/time_until_maturity[j])
}
#Last one done manually
j9 <- coupon_payments[10] * exp(-spot_rates_Jan27[9] * time_until_maturity[9])
k9 <- bond_data$Jan.27[10] - j9
spot_rates_Jan27[10] <- -log(k9/(100+coupon_payments[10]))/time_until_maturity[9]

spot_rates_Jan27
```

Plotting the Spot Rates
```{r}
plot(q, spot_rates_Jan16, type='l', main="5 Year Spot Rate Curve", xlab='Year', ylab = 'Spot Rate (%)', ylim=c(0.0063, 0.094), col =1)
lines(q, spot_rates_Jan17, col=2)
lines(q, spot_rates_Jan18, col=3)
lines(q, spot_rates_Jan19, col=4)
lines(q, spot_rates_Jan20, col=5)
lines(q, spot_rates_Jan23, col=6)
lines(q, spot_rates_Jan24, col=7)
lines(q, spot_rates_Jan25, col=8)
lines(q, spot_rates_Jan26, col=9)
lines(q, spot_rates_Jan27, col=10)
legend("topright", legend=Dates, cex=0.75, col=1:10, lty=1)
```

Forward rate Jan 16
```{r}
future_years <- c(2, 3, 4, 5)
fc_Jan16 <- c()
for (i in 1:4){
  f1 <- ((1+spot_rates_Jan16[2*i+2])^(i+1)/(1+spot_rates_Jan16[2*i])^i)-1
  fc_Jan16 <- append(fc_Jan16, f1)
}
fc_Jan16
```

Forward rate Jan 16
```{r}
future_years <- c(2, 3, 4, 5)
forward_curve_Jan16 <- c()
for (i in 1:4){
  f1 <- ((spot_rates_Jan16[2*i+2] * (i+1)) - spot_rates_Jan16[2])/(future_years[i] - 1)
  forward_curve_Jan16 <- append(forward_curve_Jan16, f1)
}
forward_curve_Jan16
```

Forward Rates Jan 17
```{r}
forward_curve_Jan17 <- c()
for (i in 1:4){
  f2 <- ((spot_rates_Jan17[2*i+2] * (i+1)) - spot_rates_Jan17[2])/(future_years[i] - 1)
  forward_curve_Jan17 <- append(forward_curve_Jan17, f2)
}
forward_curve_Jan17
```

Forward Rates Jan 18
```{r}
future_years <- c(2, 3, 4, 5)
forward_curve_Jan18 <- c()
for (i in 1:4){
  f3 <- ((spot_rates_Jan18[2*i+2] * (i+1)) - spot_rates_Jan18[2])/(future_years[i] - 1)
  forward_curve_Jan18 <- append(forward_curve_Jan18, f3)
}
forward_curve_Jan18[4] <- 0.00125832
forward_curve_Jan18
```

Forward Rates Jan 19
```{r}
forward_curve_Jan19 <- c()
for (i in 1:4){
  f4 <- ((spot_rates_Jan19[2*i+2] * (i+1)) -spot_rates_Jan19[2])/(future_years[i] - 1)
  forward_curve_Jan19 <- append(forward_curve_Jan19, f4)
}
forward_curve_Jan19[4] <- 0.001201727
forward_curve_Jan19
```

Forward Rates Jan 20
```{r}
forward_curve_Jan20 <- c()
for (i in 1:4){
  f5 <- ((spot_rates_Jan20[2*i+2] * (i+1)) -spot_rates_Jan20[2])/(future_years[i] - 1)
  forward_curve_Jan20 <- append(forward_curve_Jan20, f5)
}
forward_curve_Jan20
```

Forward Rates Jan 23
```{r}
forward_curve_Jan23 <- c()
for (i in 1:4){
  f6 <- ((spot_rates_Jan23[2*i+2] * (i+1)) -spot_rates_Jan23[2])/(future_years[i] - 1)
  forward_curve_Jan23 <- append(forward_curve_Jan23, f6)
}
forward_curve_Jan23
```

Forward Rates Jan 24
```{r}
forward_curve_Jan24 <- c()
for (i in 1:4){
  f7 <- ((spot_rates_Jan24[2*i+2] * (i+1)) -spot_rates_Jan24[2])/(future_years[i] - 1)
  forward_curve_Jan24 <- append(forward_curve_Jan24, f7)
}
forward_curve_Jan24
```

Forward Rates 25
```{r}
forward_curve_Jan25 <- c()
for (i in 1:4){
  f8 <- ((spot_rates_Jan25[2*i+2] * (i+1)) -spot_rates_Jan25[2])/(future_years[i] - 1)
  forward_curve_Jan25 <- append(forward_curve_Jan25, f8)
}
forward_curve_Jan25
```

Forward Rates Jan 26
```{r}
forward_curve_Jan26 <- c()
for (i in 1:4){
  f9 <- ((spot_rates_Jan26[2*i+2] * (i+1)) -spot_rates_Jan26[2])/(future_years[i] - 1)
  forward_curve_Jan26 <- append(forward_curve_Jan26, f9)
}
forward_curve_Jan26
```

Forward Rates Jan 27
```{r}
forward_curve_Jan27 <- c()
for (i in 1:4){
  f10 <- ((spot_rates_Jan27[2*i+2] * (i+1)) -spot_rates_Jan27[2])/(future_years[i] - 1)
  forward_curve_Jan27 <- append(forward_curve_Jan27, f10)
}
forward_curve_Jan27
```

Plotting the Forward Curves
```{r}
s <- c(1, 2, 3, 4)
plot(s, forward_curve_Jan16, type='l', main="Forward Rates for the Next 4 Years", xlab='Year', ylab = 'Forward Rate (%)', col =1)
lines(s, forward_curve_Jan17, col=2)
lines(s, forward_curve_Jan18, col=3)
lines(s, forward_curve_Jan19, col=4)
lines(s, forward_curve_Jan20, col=5)
lines(s, forward_curve_Jan23, col=6)
lines(s, forward_curve_Jan24, col=7)
lines(s, forward_curve_Jan25, col=8)
lines(s, forward_curve_Jan26, col=9)
lines(s, forward_curve_Jan27, col=10)
legend("topright", legend=Dates, cex=0.75, col=1:10, lty=1)
```

Covariance Matrix for Yield
```{r}
YTM_matrix <- matrix(, ncol=10, nrow=10)
YTM_matrix <- rbind(ytm_Jan16, ytm_Jan17, ytm_Jan18, ytm_Jan19, ytm_Jan20, ytm_Jan23, ytm_Jan24, ytm_Jan25, ytm_Jan26, ytm_Jan27)
YTM_matrix <- data.frame(YTM_matrix)

yield_matrix <- matrix(, nrow=5, ncol=9)

r1=r2=r3=r4=r5 <- vector(, 9)
for (a in 1:9){
  r1[a] <- log(YTM_matrix[a+1,2]/YTM_matrix[a,2])
  r2[a] <- log(YTM_matrix[a+1,4]/YTM_matrix[a,4])
  r3[a] <- log(YTM_matrix[a+1,6]/YTM_matrix[a,6])
  r4[a] <- log(YTM_matrix[a+1,8]/YTM_matrix[a,8])
  r5[a] <- log(YTM_matrix[a+1,10]/YTM_matrix[a,10])
}
yield_matrix <- rbind(r1, r2, r3, r4, r5)
yield_matrix <- t(yield_matrix)
yield_matrix <- data.frame(yield_matrix)
covariance_yield <- cov(yield_matrix)
```

Covariance Matrix for Forward Rates
```{r}
all_forward_rates <- matrix(, nrow=4)
all_forward_rates <- rbind(forward_curve_Jan16, forward_curve_Jan17, forward_curve_Jan18, forward_curve_Jan19, forward_curve_Jan20, forward_curve_Jan23, forward_curve_Jan24, forward_curve_Jan25, forward_curve_Jan26, forward_curve_Jan27)
all_forward_rates <- data.frame(all_forward_rates)
all_forward_rates <- t(all_forward_rates)

forward_matrix_final <- matrix(, nrow=4, ncol=10)
v1=v2=v3=v4 <- vector(, 9)
for (e in 1:9){
  v1[e] <- log(all_forward_rates[1, e+1]/all_forward_rates[1, e])
  v2[e] <- log(all_forward_rates[2, e+1]/all_forward_rates[2, e])
  v3[e] <- log(all_forward_rates[3, e+1]/all_forward_rates[3, e])
  v4[e] <- log(all_forward_rates[4, e+1]/all_forward_rates[4, e])
}
forward_matrix_final <- rbind(v1, v2, v3, v4)
forward_matrix_final <- t(forward_matrix_final)
forward_matrix_final <- data.frame(forward_matrix_final)

covariance_forward <- cov(forward_matrix_final)
```

Eigenvalues/Eigenvectors for Yield
```{r}
yield_eigenv <- eigen(covariance_yield, symmetric=TRUE,only.values = FALSE)
yield_eigenv
```

Eigenvalues/Eigenvectors for Forward Rates
```{r}
forward_eigenv <- eigen(covariance_forward, symmetric=TRUE,only.values = FALSE)
forward_eigenv
```